---
title: "RWorksheet_Gallenero#4c"
output: pdf_document
date: "2023-11-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.Use the dataset mpg
a. Show your solutions on how to import a csv file into the environment.

```{r}

library(readr)

dataset_mpg <- read.csv("mpg.csv")
dataset_mpg

```

b. Which variables from mpg dataset are categorical?

```{r}

var_mpg <- names(sapply(dataset_mpg, function(x) is.factor(x) || is.character(x)))
var_mpg

# manufacturer, model, displ, year, cyl, trans, drv, cty, hwy, fl, and class are the variables from the dataset that are categorical.

```

c. Which are continuous variables?
```{r}

cont <- sapply(dataset_mpg, is.numeric)
var_cont <- names(dataset_mpg)[cont]
var_cont

# displ, year, cyl, cty, hwy are the continous variables.

```

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
a. Group the manufacturers and find the unique models. Show your codes and result.

```{r}

library(dplyr)

Most_models <- dataset_mpg %>%
  group_by(manufacturer) %>%
  summarise(number_model = n_distinct(model)) %>%
  arrange(desc(number_model)) %>%
  head(1)


Most_variations <- dataset_mpg %>%
  group_by(model) %>%
  summarise(number_vars = n()) %>%
  arrange(desc(number_vars)) %>%
  head(1)


cat("The manufacturer with the most models is:", Most_models$manufacturer, "\n")
cat("The model with the most variations is:", Most_variations$model, "\n")


```

